### 2014-07-03 tex/main-body-text.tex
edited conclusion

Despite the similar allocation of individuals into $K$ clusters, the 
plateau analysis to estimate $K$ in {\sc flock} leads to different conclusions among groups of runs, as well as 
one that differs from {\sc structure}. While no plateaus were observed in any of the six groups 
for a $K\geq 2$, the plateau record varied greatly among each
group for $K = 2$ (plateau sequences: Run1 = 3,2,3,7,4; Run2 = 2,5,2,3,2,3; Run3 = 2,2,2,2,2,5; 
Run4 = 2,3,4,2,4,3; Run5 = 2,2,5,5,3; and; Run6 = 2,2,2,2,2,3,3). The plateau analysis 
procedure suggested by \citet{Duc&Tur2012} leads to
either an estimate of K $\geq$ 2 (Run1) or {\em undecided} as a consequence of no plateau 
lengths being greater than 6. The ($\ln P(D)$) from {\sc structure}
was largest at $K$ = 6 which was further supported by the $q_i$ plots, however,
the $\Delta K$ method supported a $K$ of 2.

@@

\section*{Conclusions}

One of the major claimed advantages of {\sc flock} over MCMC methods for solving the 
K-partition problem is that, ``The very short processing times for each run of {\sc flock} allows for
 comparing partitions from many runs of each $K$'' \citep[][p.~735]{Duc&Tur2012}.
While the authors observed improved process time in their 
comparisions we did not observe the same advantage in our analysis. In their comparison, the 
run parameters for {\sc flock} were 50 runs and 20 iterations for each $K$ from 1 to four more than 
the true value of $K$. 
These process times were compared to the admixture model with correlated allele 
frequencies in {\sc structure} with ten iterations of a 50,000 sweep burn-in followed 
by a 200,000 sweep sample from the posterior of $K$ from 1 to K+3. 
It seemed reasonable to compare, for each $K$, one iteration of {\sc structure} 
to one iteration of {\sc flock} (50 different starting allocations that are reallocated 20 times) 
as both evaluate the partition space once. The dataset used in this comparision was an order of magnatude
larger than the largest used by \citep{Duc&Tur2012} (N = 60, 120, and 240). Any process speed advantages 
may only been observed with small sample sizes.

The partition space over which {\sc flock} evaluates how individuals are clustered 
greatly increases with both the number of individuals and the number of clusters into 
which they can be partitioned. The total partition space can be 
defined by $\sum\limits_{i=1}^k S(n,k)$ where S(n,k) is a Stirling number of the 
second kind, $k$ is the number of clusters, and $n$ is the number of individuals.
The likelihood surface spanning all possible partitions
may be characterized by multiple peaks and troughs.
The peak or plateau that the  {\sc flock} algorithm finds is entirely dependent on the initial allocation of 
individuals as it has no way to visit areas of the likelihood surface with smaller values.
 It is not surprising that with 2,596 individuals we could have 
drastically different plateau sequences with just two clusters and that as $K$ increases
{\sc flock} finds suboptimal allocations. If the number of initial allocations does not increase
as $K$ is increased, the starting allocations become diffusely scattered in the space of all
possible partitions making it unlikely that they will converge.

For those that do decide to use {\sc flock} as an alternative to {\sc structure}, the adhoc rules 
to estimate K may be overly stringent. While there has been ample 
discourse concering choosing the optimal K in {\sc structure} \citep{Evannoetal2005, 
Pritchardetal2000, Wap&Gag2006} and \citet{Pritchardetal2000} urges users 
to be cautious of interpreting K, much less effort has been put forth examining how the number of individuals
and thenumber of starting allocations can affect not only the convergence of different runs, but also the 
inference of $K$. The authors developed estimation rules based on simulated data 
sets with a modest number of individuals and different migration regimes and rates, number of loci, and K. 
By relying on identical allocations (mean LLOD values) as a means of supporting 
a particular $K$,  similar but not idential 
allocations will provide no support to a given $K$.

>>>>

Despite the similar allocation of individuals into $K$ clusters, the 
plateau analysis to estimate $K$ in {\sc flock} leads to different conclusions among groups of runs, as well as 
one that differs from {\sc structure}. While no plateaus were observed in any of the six groups 
for a $K\geq 2$, the plateau record varied greatly among each
group for $K = 2$ (plateau sequences: Run1 = 3,2,3,7,4; Run2 = 2,5,2,3,2,3; Run3 = 2,2,2,2,2,5; 
Run4 = 2,3,4,2,4,3; Run5 = 2,2,5,5,3; and; Run6 = 2,2,2,2,2,3,3). The plateau analysis 
procedure suggested by \citet{Duc&Tur2012} leads to
either an estimate of K $\geq$ 2 (Run1) or {\em undecided} as a consequence of no plateau 
lengths being greater than 6. The ($\ln P(D)$) from {\sc structure}
was largest at $K$ = 6 which was further supported by the $q_i$ plots, however,
the $\Delta K$ method supported a $K$ of 3.

@@

\section*{Conclusions}
The use of genetic markers to identify population structure, specifically
the number of genetic clusters in a collection $K$ has drastically exanded
in the past decade to aid in the fields of evolution, epidemiology, and conservation
genetics. By virtute of this exanding desire to identify $K$ there has been a push to 
develop methods that can accurately infer $K$ while at the same time be easy to run
both in terms of proccess time and user interface. Naturally, a program claiming
``more accurate allocations and more reliable estimates of $K$ [while running]
much faster than {\sc structure} \citep[][p.~734]{Duc&Tur2012}'' is particularly 
appealing. Despite these two claims, we have shown that the algorithm implemented 
in {\sc flock} is equivalent to 
a restrictured case of the simulated annealing algorthim to find the Bayesian 
MAP estimate from the marginalized form of the no admixture and non-correlated
allele frequency model in {\sc structure} with no processing speed advantage for real datasets.

It is not surprising then, that because of the similaries in the algortims implemented that 
we obsered  
congruent results between the 'best run' from {\sc flock} and {\sc structure} using 
a real data set from \textit{O. mykiss}. Though we observed very similar results
between the two programs at low values of $K$, {\sc flock} had a tendency to be inconsistent
among its 'best runs' at higher values - while still finding a similar solution to {\sc structure}.
This is most likely a result of the rapidly exanding partition space that must be evaluated as $K$ increases. 
The total partition space is 
defined by $\sum\limits_{i=1}^k S(n,k)$ where S(n,k) is a Stirling number of the 
second kind, $k$ is the number of clusters, and $n$ is the number of individuals.
The likelihood surface spanning all possible partitions could
be characterized by multiple peaks and troughs.
The peak or plateau that the  {\sc flock} algorithm finds is entirely dependent on the initial allocation of 
individuals as it has no way to visit areas of the likelihood surface with smaller values.
If the number of initial allocations does not increase
as $K$ is increased, the starting allocations become diffusely scattered in the space of all
possible partitions making it unlikely that they will converge. This coupled with the fact that there is 
ambiguity in what defines a 'best run' when all plateaus are lenght zero obfuscates any interpretation
of results from {\sc flock} if no non-zero plateaus are encountered.

One of the major claimed advantages of {\sc flock} over methods for solving the 
K-partition problem is that, ``The very short processing times for each run of {\sc flock} allows for
 comparing partitions from many runs of each $K$'' \citep[][p.~735]{Duc&Tur2012}.
While the authors observed improved process time in their 
comparisions we did not observe the same advantage in our analysis. This appears to result from two issues:
(1) the number of individuals in each collection and (2) the lenght of runs for each program in each comparision. 
In their comparison, the 
run parameters for {\sc flock} were 50 runs and 20 iterations for each $K$ from 1 to four more than 
the true value of $K$. 
These process times were compared to the admixture model with correlated allele 
frequencies in {\sc structure} with ten iterations of a 50,000 sweep burn-in followed 
by a 200,000 sweep sample from the posterior of $K$ from 1 to three more than the true 
value of $K$. 
It seemed reasonable to compare, for each $K$, one iteration of {\sc structure} 
to one iteration of {\sc flock} (50 different starting allocations that are reallocated 20 times) 
as both evaluate the partition space once. We chose to run six iterations for each program
in order to compare the consistency of results between programs and among runs.
In addition to changing the number of groups of runs of {\sc flock} to compare,
the dataset used was an order of magnatude
larger than the largest used by \citep{Duc&Tur2012} (N = 60, 120, and 240). It may hold that {\sc flock}
is able to process small datasets faster than {\sc structure}, but with the rapid increase in our ability to economically
genotype a large number of individuals for a great many number of loci, this advantage may not be relevant to many 
study designs.

For those that do decide to use {\sc flock} as an alternative to {\sc structure}, the adhoc rules 
to estimate K may be overly stringent. While there has been ample 
discourse concering the choice of the optimal K in {\sc structure} \citep{Evannoetal2005, 
Pritchardetal2000, Wap&Gag2006} much less effort has been put forth examining how the number of individuals
and the number of starting allocations can affect not only the convergence of different runs, but also the 
inference of $K$. The authors developed estimation rules based on simulated data 
sets with a modest number of individuals and different migration regimes and rates, number of loci, and K. 
By relying on identical allocations (mean LLOD values) as a means of supporting 
a particular $K$,  similar but not idential 
allocations will provide no support to a given $K$. We highly advise, if using {\sc flock}, to
run multiple small batches and using the programs {\sc clumpp} \citep{Jak&Ros2007} and
{\sc district} \citep{Rosenberg2004} to visually inspect the 'best run' results.

The development of software to analyze genetic data is ever increasing. In the face of a growing
number of programs, end users must wade through a seemingly countless number of programs each with
their own assumptions, strengths, and weaknesses. Lacking a recognition of these properties
can result in the ineffectual use of both time and resources.
By providing a different prespective on the algorithm
implemented in {\sc flock} we hope to increase understanding of its applicability to a different datasets
while decreasing end user confusion and misuse.
